\chapter{Multi-vessel System State Description}
\label{chap:stateDescription}
This chapter builds on existing vessel state descriptions to formulate a multi-vessel state description. Furthermore, a novel approach to predict multi-vessel platform dynamics from individual models is proposed to  answer the research subquestion:

\begin{itemize}
	\item How can the dynamics of the multi-vessel system be represented?
\end{itemize}
The described multi-vessel state description as described in Sec. \ref{sec:notations} aims to enable consistent system notation of scenarios pertaining more than a single vessel, as mult-vessel frameworks have more objects that can be referred to. The concept and definition of a platform-coordinate system is also introduced.

For collaborative motion control of modular objects an approximate dynamical model is often desired. As model parameter estimation experiments are often infeasible for every configuration of a combined structure, a prediction of dynamical behavior is proposed using the often known dynamics of modules in Sec. \ref{platformModel}

\input{systemAnalysisNotations}

\input{multivesselplatformmodelapproach1}

\vspace{20mm}

This chapter described the multi-vessel state notation that allows consistent referencing to objects in a scenario pertaining multiple waterborn structures. Parameters that describe a parameter of a ship (such as forces, positions and velocities) are identifyable with subscripts indicating which object it pertains. The same approach has been applied to platform state description, such that various body ( $[\{b_1\},\{b_2\}, ...,\{b_i\}$), platform  ($[\{p_1\},\{p_2\}, ...,\{p_i\}$) and global (inertial frame $\{n\}$) are distinguished. 
Furthermore, an approach to predicting dynamcis of a combined waterborne structure from has been proposed that uses often known models of individual modules as building blocks. 
Both the multi-vessl state notation and the proposed platform model will be used in the next chapter during the development of a control system of the described structure. 




